# Assessing model quality {#sec-results-eval}

As mentioned in [@sec-resampling], in order to evaluate model performance nested *k*-fold cross-validation was performed.
This resulted with 2500 models created both on performance estimation level and hyperparameter tuning level.
Results of these models were then evaluated and quality measures were calculated.
In [@tbl-tabela4], overall quality measures such as accuracy and Kappa coefficient are presented.
Moreover, I calculated weighted precision, recall and F1-score.
Weights for these calculations were based on number of observations from each land cover class.
Original precision, recall and F1-score values by land cover type are shown in [@tbl-tabela5].

In general, model achieved accuracy level of 0.752 with Kappa coefficient of 0.652.
These values are rather average and model indeed needs some improvements.
On the other hand, this performance is probably enough to assess thermal band's importance ([@sec-results-therm]), which is the main goal of this study.

```{r tabela3, echo=FALSE}
#| label: tbl-tabela4
#| echo: false
#| tbl-cap: "Measures of overall model accuracy calculated during cross-validation/resampling process."
df = data.frame(
  measure = c("overall accuracy", "Kappa coefficient", "precision (user's accuracy)", "recall (producer's accuracy)", "weighted F1-score"),
  average_value = c(0.752, 0.652, 0.742, 0.751, 0.743)
)
colnames(df) = c("Measure", "Average value")
df |>
  knitr::kable(format = "latex", booktabs = TRUE, linesep = "") |> 
  kableExtra::row_spec(0, bold = TRUE, extra_css = "border-left: 1px solid grey; border-right: 1px solid grey") |>
  kableExtra::column_spec(1, border_left = TRUE) |>
  kableExtra::column_spec(2, border_left = TRUE, border_right = TRUE) |>
  kableExtra::kable_classic(full_width = FALSE) |>
  kableExtra::row_spec(1:(nrow(df) - 1), extra_latex_after = "\\hline")
```

A deeper analysis of performance measures by land cover class ([@tbl-tabela5]) shows that precision and recall values for certain type are similar.
This means that model did not have any specific problem either with too many false positive (FP) or false negative (FN) predictions.
It was just not that good in general.
Model performed very poorly in terms of correctly classifying observations of wetlands class but it is quite common issue across many studies (for example: @malinowski_automated_2020).
Also bare land class had low values of model quality with F1-score of 0.242.
The main problem concerning these land cover classes is that there was probably not enough training points for each of them.
On the other hand, two largest classes in terms of number of observations - arable land and forests - were classified much more accurately, with F1-score of 0.777 and 0.889 respectively.
Land cover type with the highest values of precision and recall, despite of low number of observations, was waterbodies class.
Model performed very good for this class probably because of its distinct spectral characteristics and easily distinguishable borders.

```{r tabela3, echo=FALSE}
#| label: tbl-tabela5
#| echo: false
#| tbl-cap: "Accuracy measures by land cover class."
df = data.frame(
  class = c("arable land", "grasslands", "forests", "bare land", "artificial land", "water bodies", "wetlands"),
  precision = c(0.732, 0.612, 0.886, 0.320, 0.656, 0.971, 0.394),
  recall = c(0.828, 0.613, 0.892, 0.194, 0.493, 1.000, 0.121),
  f1score = c(0.777, 0.612, 0.889, 0.242, 0.563, 0.985, 0.185)
)
colnames(df) = c("Land cover class", "Recall (producer's accuracy)", "Precision (user's accuracy)", "F1-score")
df |>
  knitr::kable(format = "latex", booktabs = TRUE, linesep = "") |> 
  kableExtra::row_spec(0, bold = TRUE, extra_css = "border-left: 1px solid grey; border-right: 1px solid grey") |>
  kableExtra::column_spec(1:3, border_left = TRUE) |>
  kableExtra::column_spec(4, border_left = TRUE, border_right = TRUE) |>
  kableExtra::kable_classic(full_width = FALSE) |>
  kableExtra::row_spec(1:(nrow(df) - 1), extra_latex_after = "\\hline")
```
